Hackathon SOlution 1
================================================================
        
```{r, loading_dependant_packages}
x <- c("dplyr","zoo","xlsx","readxl","bindrcpp","lubridate", "RCurl", "ggplot2")
lapply(x, require, character.only = TRUE)
```


## Setting up present working directory
```{r, pwd setting}
setwd("C:/Users/hp/datasciencecoursera/Course3(Getting and Cleaning Data)")
```

## Loading and preprocessing the data

### 1. Load the data (i.e. read.csv())
```{r, loading_data}
if(!file.exists('TriathalonAnalyticalDataSet.xls')){
        text=getURL("https://github.com/RahulPriyadarshi3785/datasciencecoursera/blob/master/Course3(Getting%20and%20Cleaning%20Data)/TriathalonAnalyticalDataSet.xls")
}

df <- tbl_df(read_excel(("TriathalonAnalyticalDataSet.xls"), sheet = 2))

```

### 2. Process/transform the data (if necessary) into a format suitable for your  analysis
```{r, data_wrangling}
## df <- tbl_df(my_data)
str(df)
```

## Answering Questions:

### 1.Which Day has most completed trips between 1st and 5th Feb?
```{r}
q1 <- df %>% select(Date, TRIPS) %>% group_by(Date) %>% summarise_all(funs(sum)) %>% head(5) %>% arrange(TRIPS)
q1
```

### 2. What hour of the day has been most completed trips on febrauary 11?

```{r}
q2 <- df %>% select(Date, Time, TRIPS) %>% filter(Date == as.POSIXlt(as.Date("2017-02-11"))) %>% group_by(Time) %>% summarise_all(funs(sum)) %>% top_n(1)
q2
```

### 3. When the No. of unique driver increases so does the number of EYEballs?
```{r}
q3 <- df %>% select(Date, EYEBALLS, UNIQUE_DRIVERS) %>% group_by(Date) %>% summarise_all(funs(sum)) %>% arrange(EYEBALLS)
q3
```

## 4. Percentage of Eyeballs occured on Friday?
```{r}
q4 <- df %>% select(Date, EYEBALLS) %>% mutate(Date = weekdays(Date)) %>% group_by(Date) %>% summarise_all(funs(sum)) %>% mutate(EYEBALLS = EYEBALLS/sum(df[,"EYEBALLS"])*100) %>% filter(Date == "Friday")
## Using ggplot2
g <- ggplot(q3, aes(x= EYEBALLS, y= UNIQUE_DRIVERS, type = Date))
g + geom_line() +
        xlab("Eyeballs") +
        ylab("Unique Drivers") +
        ggtitle('VArying Rate as the no. of Eyeballs increases')
```

### 5. HOw many hours of data are in datasets?
```{r}
q5 <- dim(df)[1]
q5
```

### 6. Weighted Request per driver for 1 days datasets?
```{r}
q6 <- apply(df[,"REQUESTS"],2, sum)/apply(df[,"UNIQUE_DRIVERS"],2, sum)
q6
```

### 7. What hour of day in 15 day in Febrauary had least number of requests?

```{r}
q7 <- df %>% select(Time, REQUESTS) %>% group_by(Time) %>% summarise_all(funs(sum)) %>% top_n(-1)
q7
```

### 8. IF you have 10 drivers in 15 day dataset when will u add them?

```{r}
q8 <- df %>% select(Time, UNIQUE_DRIVERS) %>% group_by(Time) %>% summarise_all(funs(sum)) %>% top_n(-1)
q8
```

### 9. Use data to detemine when the best day time is for drivers?

```{r}
q9 <- df %>% select(Time, REQUESTS, UNIQUE_DRIVERS) %>% group_by(Time) %>% summarise_all(funs(sum)) %>% mutate(Best_TIme = REQUESTS/UNIQUE_DRIVERS) %>% arrange(Best_TIme) %>% head(1)
q9
```
##YOu need to be careful while analysing data in this question


### 10. Name the 72 hours of period with highest an=mouts of zeroes.
```{r}
##YOu need to be careful while analysing data in this question
q10 <- df %>% select(Time, UNIQUE_DRIVERS) %>% filter(UNIQUE_DRIVERS != 0) %>% group_by(Time) %>% summarise(count = n()) %>% tail(15)
q10
## Answer is 7 to 10
```


### 11. If 15 days dataset has 8 hours shift then which of the shift has maximum numbers of Requests

```{r}
q11 <- df %>% select(Time, REQUESTS) %>% group_by(Time) %>% summarise_all(funs(sum)) %>% mutate(val = rollsumr(REQUESTS, k = 8, fill = NA)) %>% filter(row_number()%%8 == 0) %>% top_n(1)
q11
```

### 12. On which day the TRIPS/REQUESTS is highest?
```{r}
q12 <- df %>% select(Date, TRIPS, REQUESTS) %>% group_by(Date) %>% summarise_all(funs(sum)) %>% mutate(TRIPs_PER_REQUEST = TRIPS/REQUESTS) %>% arrange(TRIPs_PER_REQUEST) %>% tail(1)
q12
```